### 2021-05-08 manuscript.tex
performed spellcheck

\section{Abstract}
Maintenance and upgrades are a fundamental requirement of any software-based tool, and the resources within the field of Bioinformatics are no exception. Too often, the combination of using self-taught programmers who are typically trained Biologists along with the broader lack of a software development ecosystem can propagate elevated degrees of software entropy, also known as technical debt. It is then a fair assumption to say that this level of entropy is higher than in fields which more consistently employ the skills of professional developers. Here, we present the case of the Candidate Cancer Gene Database (CCGD, https://z.umn.edu/ccgd) as a minor instance of taking advantage of an upgrade cycle in an attempt to service some of this technical debt. The goals of this project ultimately strived to acheive the implementation of a new, core framework at the base of the data integration backend as well as a more modern and simplified web front-end. In all, this represented a complete rebuild of the application while preserving the existing conceptual implementation. Our work served to establish a new application with greater structural simplicity, improved resource use, and, most relevant, a code base which will be easier to maintain and upgrade in the future.

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/rmarkdown_flow.pdf}
    \caption[Rmarkdown Rendering Flow]{\textbf{Rmarkdown Rendering Flow} The core web framework was created using a stitch process composed of two document conversion steps, which are both contained in the single \texttt{rmarkdown::render\_site} command. These two transformations are done by the R-specific knitr package and then the open-source Pandoc tool. The first phase holds less relevance in particular as its role is to execute R code chunks, of which there are few in our page content. The markdown output from this step is seamlessly injected into a Pandoc conversion which offers a large variety in document output. Specific to our case, the Pandoc conversion paired with a companion YAML configuration file allows for multidocument markdown-to-html rendering which is assembled as a complete site stack. As a single process, this function generates an integrated, modern web front-end from content written in a light-weight syntax with almost none of the typical complexity of full-scale web page creation.}\label{fig:renderFlow}
\end{figure}

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/file_tree.pdf}
    \caption[Project Directory File Tree Comparison]{\textbf{Project Directory File Tree Comparison} The above figure offers a qualitative assessment of an element of technical debt that was improved in refactoring. Shown are the horizontally oriented file trees of the site directories of the new and old projects. What is illustrated is a dramatic improvement in the tree size and complexity through the elimination of a large degree of redudant or unecessary content, log, and style files. This is accentuated even further when comparing the source code. This was accomplished both in the intentional design of eliminating content redundancy and in the streamlined structure of the Rmarkdown rendering process. This allowed for the elimination of several dimensions of broader web design and table build data elements.}\label{fig:fileTree}
\end{figure}

>>>>

\section{Abstract}
Maintenance and upgrades are a fundamental requirement of any software-based tool, and the resources within the field of Bioinformatics are no exception. Too often, the combination of using self-taught programmers who are typically trained Biologists along with the broader lack of a software development ecosystem can propagate elevated degrees of software entropy, also known as technical debt. It is then a fair assumption to say that this level of entropy is higher than in fields which more consistently employ the skills of professional developers. Here, we present the case of the Candidate Cancer Gene Database (CCGD, https://z.umn.edu/ccgd) as a minor instance of taking advantage of an upgrade cycle in an attempt to service some of this technical debt. The goals of this project ultimately strived to achieve the implementation of a new, core framework at the base of the data integration backend as well as a more modern and simplified web front-end. In all, this represented a complete rebuild of the application while preserving the existing conceptual implementation. Our work served to establish a new application with greater structural simplicity, improved resource use, and, most relevant, a code base which will be easier to maintain and upgrade in the future.

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/rmarkdown_flow.pdf}
    \caption[Rmarkdown Rendering Flow]{\textbf{Rmarkdown Rendering Flow} The core web framework was created using a stitch process composed of two document conversion steps, which are both contained in the single \texttt{rmarkdown::render\_site} command. These two transformations are done by the R-specific knitr package and then the open-source Pandoc tool. The first phase holds less relevance in particular as its role is to execute R code chunks, of which there are few in our page content. The markdown output from this step is seamlessly injected into a Pandoc conversion which offers a large variety in document output. Specific to our case, the Pandoc conversion paired with a companion YAML configuration file allows for multi-document markdown-to-html rendering which is assembled as a complete site stack. As a single process, this function generates an integrated, modern web front-end from content written in a light-weight syntax with almost none of the typical complexity of full-scale web page creation.}\label{fig:renderFlow}
\end{figure}

@@

\begin{figure}[H]
    \center{}
    \includegraphics[width=0.8\textwidth]{fig/file_tree.pdf}
    \caption[Project Directory File Tree Comparison]{\textbf{Project Directory File Tree Comparison} The above figure offers a qualitative assessment of an element of technical debt that was improved in refactoring. Shown are the horizontally oriented file trees of the site directories of the new and old projects. What is illustrated is a dramatic improvement in the tree size and complexity through the elimination of a large degree of redundant or unnecessary content, log, and style files. This is accentuated even further when comparing the source code. This was accomplished both in the intentional design of eliminating content redundancy and in the streamlined structure of the Rmarkdown rendering process. This allowed for the elimination of several dimensions of broader web design and table build data elements.}\label{fig:fileTree}
\end{figure}
